\textbf{Advection versus diffusion in brain intraventricular transport}

### 2025-03-27 paper/main.tex
fix ref

\textbf{Cardiac-induced/no-cilia model.} For this model version,
the only physical mechanism inducing fluid flow is the cardiac cycle.
The cardiac cycle generates directional back-and-forth motion of the CSF
in the zebrafish brain ventricles.
We model the influence of the cardiac cycle with a normal pressure boundary condition:
\begin{subequations}
    \begin{alignat}{2}
      \uu\cdot\nn = 0, \ \hat{\bsig}_{\parallel} &= \mathbf{0} \quad &&\mathrm{on} \ \Gs,
      \label{eq:modelB_slip_BC} \\
      (\mu\nabla\uu - p\mathbf{I})\nn &= \Tilde{p}(t)\nn \quad &&\mathrm{on} \ \Gp,
      \label{eq:modelB_pressure_BC}
    \end{alignat}%
\end{subequations}%
where the boundary $\Gamma = \Gs\cup\Gp$ consists of a free-slip boundary $\Gs$,
and a part $\Gp$ where a normal traction type of boundary condition is imposed.
Note, however, that the transpose velocity gradient term must be
included inside the parentheses on the left-hand side of~\Cref{eq:modelB_pressure_BC}
for it to be the traction $\bsig$. Since this is not the case,
we rather refer to this boundary condition as a normal pressure boundary condition.
The boundary $\Gs$ is handled as discussed in the previous section,
such that the stress boundary integral in~\Cref{eq:boundary_integral_BC1}
vanishes over $\Gs$. Therefore, we now have
\begin{align}
    \intG{\hat{\bsig}\cdot\vv} = \intGp{\hat{\bsig}\cdot\vv}
    &= \intGp{(\mu(\nabla\uu + (\nabla\uu)^T) - p\mathbf{I})\nn\cdot\vv}\notag \\
    &= \intGp{\mu(\nabla\uu)^T\nn\cdot\vv} + \intGp{(\mu\nabla\uu
    - p\mathbf{I})\nn\cdot\vv}\notag\\
    &= \intGp{\mu(\nabla\uu)^T\nn\cdot\vv} + \intGp{\Tilde{p}(t)(\vv\cdot\nn)}.
    \label{eq:modelB_boundary_integral1}
\end{align}
We leave the first term on the left-hand side
of~\Cref{eq:modelB_boundary_integral1} in the weak form of
the momentum equation. The second term is determined by
the boundary condition defined in~\Cref{eq:modelB_pressure_BC},
and is therefore moved to the right-hand side of the weak form.
The weak form of the momentum equation is then
\begin{equation}
    \begin{aligned}
    \intO{\rho\pdifft{\uu}\cdot\vv} + \intO{2\mu\beps(\uu) : \beps(\vv)} -
    \intO{p\nabla\cdot\vv} &-\intGp{\mu(\nabla\uu)^T\nn\cdot\vv} \\
    &=\intGp{\Tilde{p}(t)(\vv\cdot\nn)}, \quad\forall\vv\in\VV.
    \end{aligned}
    \label{eq:weak_form_mom_eq_modelB}
\end{equation}
Also considering the weak form of the continuity equation,
the weak formulation for the cardiac-only/no-cilia model version
then reads: find $(\uu,\, p)\in\VV\times Q$
such that~\Cref{eq:weak_form_mom_eq_modelB} and~\Cref{eq:pressure_weak_form}
hold for all $(\vv, \, q)\in\VV\times Q$.
To simulate the back-and-forth motion induced by the cardiac cycle observed in
previous work~\cite{Olstad2019CiliaryDevelopment}, we apply sinusoidal forcing.
The boundary $\Gp$ has two separate regions: one located on the
anterior ventricle and one on the posterior ventricle~(\Crefpanels{fig:fig1}{d}, blue markers).
For the normal pressure boundary condition, we set $\Tilde{p}(t) = 0$ on the anterior part and
$\Tilde{p}(t) = p_c(t) = -A\sin{\omega t}$ on the posterior part. The value of the amplitude
$A$ was based on heuristics and $\omega = 2\pi f$, where $f=2.22$ Hz
is the cardiac cycle frequency.

>>>>

\textbf{Cardiac-induced/no-cilia model.} For this model version,
the only physical mechanism inducing fluid flow is the cardiac cycle.
The cardiac cycle generates directional back-and-forth motion of the CSF
in the zebrafish brain ventricles.
We model the influence of the cardiac cycle with a normal pressure boundary condition:
\begin{subequations}
    \begin{alignat}{2}
      \uu\cdot\nn = 0, \ \hat{\bsig}_{\parallel} &= \mathbf{0} \quad &&\mathrm{on} \ \Gs,
      \label{eq:modelB_slip_BC} \\
      (\mu\nabla\uu - p\mathbf{I})\nn &= \Tilde{p}(t)\nn \quad &&\mathrm{on} \ \Gp,
      \label{eq:modelB_pressure_BC}
    \end{alignat}%
\end{subequations}%
where the boundary $\Gamma = \Gs\cup\Gp$ consists of a free-slip boundary $\Gs$,
and a part $\Gp$ where a normal traction type of boundary condition is imposed.
Note, however, that the transpose velocity gradient term must be
included inside the parentheses on the left-hand side of~\Cref{eq:modelB_pressure_BC}
for the expression to be the stress tensor $\bsig$, thus the left-hand side of
\Cref{eq:modelB_pressure_BC} does not constitute the normal traction $\bsigperp$.
Because of this, we rather refer to this boundary condition
as a normal pressure boundary condition.
The boundary $\Gs$ is handled as discussed in the previous section,
such that the stress boundary integral in~\Cref{eq:boundary_integral_BC1}
vanishes over $\Gs$. Therefore, we now have
\begin{align}
    \intG{\hat{\bsig}\cdot\vv} = \intGp{\hat{\bsig}\cdot\vv}
    &= \intGp{(\mu(\nabla\uu + (\nabla\uu)^T) - p\mathbf{I})\nn\cdot\vv}\notag \\
    &= \intGp{\mu(\nabla\uu)^T\nn\cdot\vv} + \intGp{(\mu\nabla\uu
    - p\mathbf{I})\nn\cdot\vv}\notag\\
    &= \intGp{\mu(\nabla\uu)^T\nn\cdot\vv} + \intGp{\Tilde{p}(t)(\vv\cdot\nn)}.
    \label{eq:modelB_boundary_integral1}
\end{align}
We leave the first term on the left-hand side
of~\Cref{eq:modelB_boundary_integral1} in the weak form of
the momentum equation. The second term is determined by
the boundary condition defined in~\Cref{eq:modelB_pressure_BC},
and is therefore moved to the right-hand side of the weak form.
The weak form of the momentum equation is then
\begin{equation}
    \begin{aligned}
    \intO{\rho\pdifft{\uu}\cdot\vv} + \intO{2\mu\beps(\uu) : \beps(\vv)} -
    \intO{p\nabla\cdot\vv} &-\intGp{\mu(\nabla\uu)^T\nn\cdot\vv} \\
    &=\intGp{\Tilde{p}(t)(\vv\cdot\nn)}, \quad\forall\vv\in\VV.
    \end{aligned}
    \label{eq:weak_form_mom_eq_modelB}
\end{equation}
Also considering the weak form of the continuity equation,
the weak formulation for the cardiac-only/no-cilia model version
then reads: find $(\uu,\, p)\in\VV\times Q$
such that~\Cref{eq:weak_form_mom_eq_modelB} and~\Cref{eq:pressure_weak_form}
hold for all $(\vv, \, q)\in\VV\times Q$.
To simulate the back-and-forth motion induced by the cardiac cycle observed in
previous work~\cite{Olstad2019CiliaryDevelopment}, we apply sinusoidal forcing.
The boundary $\Gp$ has two separate regions: one located on the
anterior ventricle and one on the posterior ventricle~(\Crefpanels{fig:fig1}{d}, blue markers).
For the normal pressure boundary condition, we set $\Tilde{p}(t) = 0$ on the anterior part and
$\Tilde{p}(t) = p_c(t) = -A\sin{\omega t}$ on the posterior part. The value of the amplitude
$A$ was based on heuristics and $\omega = 2\pi f$, where $f=2.22$ Hz
is the cardiac cycle frequency.
